New lower bounds for Schur and weak Schur numbers

### 2021-09-30 article/schur.tex
modifs (nom nul mais pas d'inspi ¯\_(ツ)_/¯)

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b.\]
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened and as a result, we can change the coloring on the last row to extend the previous partitions.
\newline

>>>>

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - q S(n))\), we have
\[ S(n+k) \geqslant q S(k) + b.\]
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened because 
the constraints in the definition of a S-template prevent a number from both being the sum of two of numbers of its own color 
as well as creating a sum in its own color, the later of which is not always relevant on the last row given that it is not completely filled.
As a result, we can change the coloring on \([\![ q S(k) + 1,  q S(k) + b]\!]\) (the last row) to extend the previous partitions.
\newline

### 2021-09-30 article/weakschur.tex
modifs (nom nul mais pas d'inspi ¯\_(ツ)_/¯)

\begin{proposition}
\label{prop7}
Let \(x, y \in \mathbb{Z}\) such that \(\pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!]\). Then
\[
\lambda(x + y) = \lambda(x) + \lambda(y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
By Proposition \ref{prop5}, \(x + y = a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a\). Then
\[
\begin{array}{l l l}
	\lambda(x + y) & = & \left \lfloor \dfrac{x + y - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \left \lfloor \dfrac{a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \lambda(x) + \lambda(y) - 1 \left \lfloor \dfrac{\pi(x) + \pi(y) - b - 1}{a}\right\rfloor \\
	 & = & \lambda(x)+\lambda(y) -1 +1 ~~~~\text{since } \pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!] \\
	 & =& \lambda(x)+\lambda(y)
\end{array}
\]
\end{proof}

@@

Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

@@

With \( S(9) \geqslant 17\,803 \), we found a new lower bound for \(\WS(10)\) using (\ref{WS(n+1)}).
Moreover, inequality (\ref{WS(n+3)}) gives new lower bounds for \(\WS(9)\), \(\WS(10)\), \(\WS(14)\) and \(\WS(15)\).

@@

The weak Schur partitions obtained with WS-templates have an extremely regular structure. Therefore, one may 
expect to find larger partitions by allowing more freedom in the structure of the partition while still preserving a structure 
somewhat close to a template-based partition thus reducing the search space to a manageable size.

@@

However, trying out all of the 2\,447\,113\,088 Schur number 5-partitions (\cite{Heule2017}) one by one would not result in a 
reasonable computation time: it is necessary to test the construction on several partitions at once. We encoded the problem as a 
satisfiability problem and used the 1616 backdoors that were used in \cite{Heule2017} in order to encode a group of partitions in 
a compact and efficient way. Among all of the 1616 backdoors, only the 911\textsuperscript{th} backdoor gave a weakly sum-free 
partition of length 643. This backdoor gave a weakly sum-free partition of length 646 and cannot give a weakly sum-free 
partition of length 647. A weakly sum-free partition of length 646 can be found in the \hyperref[WS(6)]{appendix}.

@@

We started by giving a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then generalized this construction by introducing WS-templates. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates. We reckon better WS-templates with \(n \geqslant 5\) colors can be found by making different assumptions 
and using a different method (Monte-Carlo methods for instance).

>>>>

\begin{proposition}
\label{prop7}
Let \(x, y \in \mathbb{Z}\) such that \(\pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!]\). Then
\[
\lambda(x + y) = \lambda(x) + \lambda(y).
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
By Proposition \ref{prop5}, \(x + y = a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a\). Then
\[
\begin{array}{l l l}
	\lambda(x + y) & = & \left \lfloor \dfrac{x + y - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \left \lfloor \dfrac{a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \lambda(x) + \lambda(y) - 1 \left \lfloor \dfrac{\pi(x) + \pi(y) - b - 1}{a}\right\rfloor \\
	 & = & \lambda(x)+\lambda(y) -1 +1 ~~~~\text{since } \pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!] \\
	 & =& \lambda(x)+\lambda(y).
\end{array}
\]
\end{proof}

@@

Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
with a different method,for instance with Monte-Carlo methods as in \cite{Bouzy2015AnAP}. This could be the suject of a 
future work.

@@

@@

The weak Schur partitions obtained with WS-templates have an extremely regular structure. Therefore, one may 
expect to find larger partitions by allowing more freedom in the structure of the partitions while still preserving a structure 
somewhat close to a template-based partition thus reducing the search space to a manageable size.

@@

However, trying out all of the 2\,447\,113\,088 Schur number 5-partitions \cite{Heule2017} one by one would not result in a 
reasonable computation time: it is necessary to test the construction on several partitions at once. We encoded the problem as a 
satisfiability problem and used the 1616 backdoors that were used in \cite{Heule2017} in order to encode a group of partitions in 
a compact and efficient way. Among all of the 1616 backdoors, only the 911\textsuperscript{th} backdoor gave a weakly sum-free 
partition of length 643. This backdoor gave a weakly sum-free partition of length 646 and cannot give a weakly sum-free 
partition of length 647. A weakly sum-free partition of length 646 can be found in the \hyperref[WS(6)]{appendix}.

@@

We started by giving a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then generalized this construction by introducing WS-templates. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates; thus questionning the methods used and the assumptions made when obtaining these former lower 
bounds. We reckon better WS-templates with \(n \geqslant 5\) colors can be found by making different assumptions 
and using a different method (Monte-Carlo methods for instance).
